PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-02-17 txt/manuscript.tex
?
reading again and add some cite - tommorrow add more cite and check appendix last time

\author{Paulina Lewandowska}
\author{Profesor Kierownik \L ukasz Pawela}

@@

Noisy Intermediate-Scale Quantum (NISQ)~\cite{preskill} devices are storming the market,
with a wide selection of devices based on different architectures and accompanying software
solutions. Among hardware providers offering public access to their gate--based devices, one could
mention Rigetti, IBM, Oxford Quantum Group, IonQ or Xanadu. Other vendors offer devices operating in
different paradigms. Notably, one could mention D-Wave and their quantum
annealers, or QuEra devices based on neural atoms. \todo[inline]{Expand this list and provide short
description of those architectures}. Most vendors provide their own software stack and
application programming interface for accessing their devices. To name a few, Rigetti's computers
are available through their Forest SDK and PyQuil library and IBM Q computers can be accessed
through Qiskit or IBM Quantum Experience web interface. Some cloud services, like Amazon Braket, offer access to several quantum devices under a unified API. On top of that, several
libraries and frameworks can integrate with different hardware vendors. Examples of such frameworks
include IBM Q's Qiskit or Zapata Computing's Orquestra.

It is well known that NISQ devices have their limitations. The question is to what extent those
devices can perform meaningful computations? To answer this question, one has to devise a
methodology for benchmarking them. For gate--based computers, on which this paper focuses, there
already exist several approaches. \todo[inline]{List the methods}

\begin{itemize}
	\item several examples of benchmarks -- all our metrics are readily available for devices with two qubits only and circuit depths $< 10$  for all benchmarks it should be possible to execute
	tweaked variants on larger devices when they become available \cite{cornelissen2021scalable}
	\item random benchmarking \cite{liu2022sampling, knill2007randomized, wallman2014randomized} and general overwiev see here \cite{helsen2022general}
	\item MQT Bench: Benchmarking Software and
	Design Automation Tools for Quantum Computing \cite{mqt2022} -- przejebane 
	\item QUARK framework for quantum compjuting application benchmarking for QUBO using D-QAwe ? \cite{quark2022} --  (e. g., a novel QUBO formulation
	of the partial MAX-SAT problem) and benchmark these on
	different infrastructures (e. g., D-Wave and simulation). 
	\item SupermarQ: A Scalable Quantum Benchmark Suite -- Recent works within the quantum computer architecture
	community have taken the first steps towards quantum
	benchmarking. The PPL+2020 [16] suite was evaluated on
	seven superconducting QPUs, focused on characterizing the
	error rates of different operations, and demonstrated the
	time dependence of their performance. The TriQ [17] suite
	was used to perform a cross-platform comparison between
	superconducting and trapped-ion systems and revealed the
	importance of software visibility into the hardware’s native
	gates. However, the scalability of these suites is limited by
	their reliance on circuit simulation to estimate how well the
	QPUs are performing. SupermarQ extends these works by
	introducing a systematic and principled approach to building
	a scalable quantum benchmark suite \cite{supermarq2022}

\end{itemize}
\todo[inline]{What benchmarking techniques are readily available?}

In this paper, we introduce yet another method for benchmarking gate--based devices with a simple
operational interpretation. In our method, we test how well the given device can distinguish between
two von Neumann measurements performed in different bases. We implemented our method in an
open-source Python library called PyQBench. The library supports any device available through Qiskit
library, and thus can be used with providers such as IBM Q or Amazon Braket. Along with the library,
the PyQBench package contains a command line tool for running benchmarks in most common scenarios.

\begin{itemize}
	\item NISQs are becoming widespread and easily available for the general audience
	\item As with any computational technology, there's a need for benchmarking NISQs to assess
	their potential usability.
	\item Our work aims to provide a new benchmarking scheme for gate-based devices.
	\item Why is PyQBench important?
	\begin{itemize}
		\item There exist multiple benchmarking methods already on the market
		\item Most of them have the following drawbacks:
		\begin{itemize}
			\item Their interpretation is not very intuitive (We cannot phrase it like this without
			explaining why it is so)
			\item Hard to use (is this true?)
			\item No libraries are readily available in Python (needs to be checked).
		\end{itemize}
		\item PyQBench is easy to install, it works with any Qiskit backend, and its results are
		easy to understand and interpret.
	\end{itemize}
\end{itemize}

\todo[inline]{In particular, we compile circuits to standard gate sets of the vendor using compiler RECZNEGO?, and we perform experiments across different qubit subsets.}

quantum volume - '' has become the de-facto standard benchmark to quantify the capability of Noisy Intermediate-Scale Quantum (NISQ) devices '' \cite{pelofske2022volume}

Noisy intermediate-scale quantum
(NISQ)~\cite{preskill} devices are currently storming the market. We
have a selection of readily available quantum computers based on different
architectures. Historically, the first widely introduced quantum computing
architecture is D-Wave's quantum annealer.

Additional approaches to quantum computing are currently in development. For
instance we have Ionq's computer based on ion traps and Xanadu's optical
lattices. As they are in prototype versions no access to general public is
provided.

Next, we have computers implementing the gate model of quantum computation.
These are the best potentially developed machines, which can be thought
of as fully quantum computers, by which we understand that the qubits can be in
an entangled state.

Currently one of the main providers of such architectures is Rigetti and its
Quantum Cloud Services platform. The company has
released three generations of QPUs so far: Rigetti 8Q Agave, Rigetti 19Q Acorn
and Rigetti 16Q Aspen-1. The most recent one was deployed in November 2018 and
is a downgrade in terms of number of qubits compared to the previous 19 qubit
one. However, the new chip is reportedly much more robust to noise and the
proposed architecture is much easier to scale up. The new unit is supposed to
serve as a building block for a 128 qubit
system~\footnote{https://medium.com/rigetti/the-rigetti-128-qubit-chip-and-what-it-means-for-quantum-df757d1b71ea}.
As was the case for IBM, Rigetti also provides a \texttt{Python} library,
\texttt{pyquil}~\cite{}, which enables easy access to the machine.

\todo[inline]{co napisac o tym Amazonie, czy nic ?}

\todo[inline]{rysunki architektur}

%
%\begin{figure}[h]
%	\centering
%
%	\includegraphics[width=0.9\linewidth]{aspen.png}
%	\caption{?}
%	\label{fig:aspen}
%\end{figure}

Now the natural question arises: how to construct a good metric of the
computational power of such devices?
Lately, there have emerged propositions from the scientific community on how to
benchmark such devices.

 First let us mention the work by Michielsen
\emph{et.al.}~\cite{michielsen2017benchmarking}. In there the authors study the
IBM-QE device. Their first test is the creation of the entangled state
$\frac{1}{\sqrt{2}} (\ket{01} - \ket{10})$. Next they implement a
two-qubit+two-qubit adder. Another test is an identity operation, realized by an
even number of CNOT operations. Their final test is a quantum error correction
scheme (is it possible with measurement-controlled-operations?). The final
conclusion is that, except for simple circuits, the device fails to return
correct results.

Another, recent approach is utilizing quantum communication protocols in testing
of quantum architectures~\cite{zhukov2019quantum}. Here, the authors decide to
utilize superdense coding and the BB84 quantum cryptography protocol as
benchmarks for the IBM Q devices. They utilize the mutual information of the
transferred bits as a figure of merit for the quantum computer.

Other approaches focus on generative model training. One of these
works~\cite{hamilton2018generative} is aimed at small architectures, up to 5
qubits. Another recently presented is much more
robust~\cite{benedetti2018generative}. The main drawback of both of them is the
level of complication.

The goal of this work is to introduce a new benchmark for Rigetti
architecture...

\label{}
\todo[inline]{
	Introduce the scientific background and the motivation for developing the
	software. - motywacje

	Explain why the software is important, and describe the exact (scientific)
	problem(s) it solves. - czemu ejst wazne, jakie problemy rozwiazuje

	Indicate in what way the software has contributed (or how it will
	contribute in the future) to the process of scientific discovery; if
	available, this is to be supported by citing a research paper using the
	software. - do czego sie to przyczyni w przyszlosci

	Provide a description of the experimental setting (how does the user use
	the software?). - w jaki sposób użytkownik korzysta z oprogramowania?

	Introduce related work in literature (cite or list algorithms used, other
	software etc.). - inne oprogramowania i algorytmy }

!!!!!!!

Suppose we have access to a device performing one of the predefined von Neumann
measurements, $\PP$ or $\QQ$. In PyQBench, we restrict ourselves to discriminating von Neumann measurements.
This is because, unlike other measurement types, they can be implemented on actual hardware. While the $\PP$ and $\QQ$ are known, it is not known which of them is
performed when the device executes. Based on the measurement outcome, you have to guess which
measurement was performed (but you can perform arbitrary unitary operations before and after the
measurement). What is the highest probability of making a correct guess? And what do you have to do
to achieve it? And, most importantly, why would we want to do this?

Suppose that you know a strategy that, for an ideal device, would yield a probability
$p_{\text{succ}}$ of successfully discriminating between two measurements. Will the probability be
the same on an actual physical device? For current, Noisy Intermediate Scale Quantum devices
(NISQs), the answer is: certainly not. However, the error rate that you make when guessing can be
used as a benchmarking metric. PyQBench helps you in organizing such discrimination experiments for
a single qubit system, executing them on real hardware or simulators, and
computing discrimination probabilities based on the measured bitstrings.

@@

In general, the discrimination scheme, presented in Fig.~\ref{fig:theoretical_scheme}, requires an
auxiliary system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, another measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise
we guess that it was $\mathcal{P}_\Id$.

@@

Current NISQ devices are unable to perform conditional measurements, which is the biggest
obstacle to implementing our scheme on real hardware. Luckily, we can overcome it by
cleverly adjusting our scheme so that it only uses components available on current devices.
We have two possible choices here: using a postselection or a direct sum
$V_0^\dagger\oplus V_1^\dagger$.

@@

	The first idea uses a postselection. Instead of performing a conditional measurement, for each
	measurement $\PP_U, \PP_\Id$ to be discriminated and each choice of $k \in \{0, 1\}$ we run circuit presented in Fig. \ref{fig:postselection}.

@@

		\caption{
			A schematic representation of the setup for distinguishing
			measurements $\PP_{U}$ and $\PP_{\Id}$ using direct sum.
		}\label{fig:postselection}
	\end{figure}

%	But now our experiment does not match the previously described scheme, right?
%	Yes, unless we discard all the outcomes for which $i\ne k$ (hence the name \emph{postselection},
%	we are selecting valid outcomes after the experiment is performed).

	% \begin{scheme}(Postselection)

	% %\begin{figure}[h!]
	% %	\includegraphics[scale=1.5]{onequbit.pdf}
	% %	\caption{A schematic representation of the setup for distinguishing
	% %		measurements using postselection.}
	% %	\label{postselection}
	% %\end{figure}

	% \begin{enumerate}
	% \item We prepare  the discriminator $\ket{\psi_{0}}$ on bipartite system.
	% \item We perform one of two quantum channels, either $\Phi_{\Id}$ or
	% $\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left( \Phi \otimes \Id\right)(\proj{\psi_0})$.
	% \item We measure the first part of the input state  $\ket{\psi_{0}}$ and
	% receive an output label $i$.
	% \item
	% Based on Holevo--Helstrom theorem, we perform a conditional binary
	% measurement	$\PP_{V_j}$ and hence, we implement the quantum channel $\Phi_{V_j^\dagger}$ on the second part.
	% \item To calculate the probability of correct discrimination, we include just those cases for which $i = j$.
	% \item We measure the second part of the input state  $\ket{\psi_{0}}$ and make a
	% decision based on received label $j$. If $i=j=0$, then we decide that
	% $\PP_U$ occurs.  If $i=j=1$, we decide that
	% $\PP_{\Id}$ occurs.
	% \end{enumerate}
The experiments can be grouped into classes identified by tuples of the form
	$(\mathcal{Q}, k, i, j)$, where $\mathcal{Q} \in \{\PP_U, \PP_\Id\}$ denotes the chosen measurement, $k \in \{0,1\}$, $i \in \{0,1\}$ and $j \in \{0,1\}$.
	In postselection approach, we discard all the experiments for which $k \ne i$. Hence, the total number of valid experiments is

@@

	Finally, we count the experiments (among the valid ones) resulting in successful discrimination.
	If we have chosen $\PP_U$, then we guess correctly iff $j=0$. Similarly, for
	$P_\Id$, we guess correctly iff $j=1$. If we define
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_U, k = i, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_\Id, k = i, j = 1\},
	\end{eqnarray}
	then the empirical success probability can be computed as

@@

The second idea uses  the direct sum $V_0^\dagger \oplus V_1^\dagger.$    Here, instead of performing a conditional
measurement $\PP_{V_k}$, where $k\in \{0,1\}$ we run circuit presented in	Fig.~\ref{fig:controlled}.
%	\begin{enumerate}
%		\item We prepare the discriminator $\ket{\psi_{0}}$ on bipartite
%		system.
%		%\item We perform one of two quantum channels, either $\Phi_{\Id}$ or
%		%$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left(\Phi \otimes \Id\right)(\proj{\psi_0})$.
%		\item We apply one of unitary,   either $U^\dagger$ or
%		$\Id$, on the first system.
%		%\item Based on Holevo--Helstrom theorem, we perform a conditional binary measurement $\PP_{V_j}$, and hence, we implement the direct sum of quantum channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$. Observe that   $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger} = \proj{0} \otimes \Phi_{V_0^\dagger}  + \proj{1} \otimes \Phi_{V_1^\dagger}.  $
%		\item We apply direct sum $V_0^\dagger \oplus V_1^\dagger$ on the whole systems.
%		\item We measure both systems in computational basis $\Delta$.
%		\item We make a decision based on the received label $j$ on the second system. If $j=0$, then we
%		decide that $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs.
%	\end{enumerate}

%	The schematic representation of this setup is depicted in
%	Fig.~\ref{fig:controlled}.
	\begin{figure}[h!]
		\centering
		\includegraphics[scale=1.5]{pics/controlled_unitary}

@@

So far, we only discussed how the discrimination is performed using two different
schemes, assuming that all needed components ($\ket{\psi_0}$, $V_0$, $V_1$) are known. It turned out that the determination of $V_0$ and $V_1$ is a simple task while the discriminator $\ket{\psi_{0}}$ is known. However,
sad to say, determining the analytical form of the discriminator $\ket{\psi_{0}}$
is, in general, NP-hard problem.
We have found the analytical solution of the discrimination scheme for a particular class of von Neumann measurements.

@@

\begin{equation}
U = H
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this case, the optimal input state is Bell state of the form
\begin{equation}
\ket{\psi_{0}} = \frac{1}{\sqrt{2}} \ket{00} + \ket{11}.
\end{equation}
And, for a given angle  $\phi \in  [0, 2 \pi)$,   the unitaries $V_0$,  $V_1$
have the following form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi -
	\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right)
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
Finally, we have also calculated the theoretical probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ is given by
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} .
\end{equation}
More details and proofs of the construction's validity   we can find in the \ref{app:optimal-probability} and \ref{app:optimal-strategy}.

@@

\section{PyQBench as a library}\label{sec:pyqbench-as-library}
This guide introduces basic concepts needed for using PyQBench as a library from a Python script. In particular, we will cover the following topics:
\begin{enumerate}
\item Defining experiment using circuit components.
\item Assembling circuits needed for given benchmarking scheme.
\item Running experiment on a simulator or a hardware.
\item Obtaining empirical probability of successful discrimination between measurements.
\end{enumerate}

On the conceptual level, every discrimination experiment performed in PyQBench needs the following:
\begin{enumerate}
\item Discriminator, i.e. optimal initial state for the circuit.
\item Unitary $U^\dagger$ defining alternative measurement to be discriminated from the measurement in Z-basis.
\item $V_0^\dagger$ and $V_1^\dagger$, positive and negative parts of optimal measurement from Holevo-Helstrom theorem.
\end{enumerate}
%Currently available quantum computers typically cannot start execution from an arbitrary state. Instead, we have to prepare it ourselves. Hence, to execute experiment using postselection scheme, we need an instruction taking to the desired discriminator $\ket{\psi_0}$
%and instructions implementing $U^\dagger$, $V_0^\dagger$ and $V_1^\dagger$. Whereas, execute experiment using direct sum approach, we need an instruction taking to the desired discriminator  $\ket{\psi_0}$
%and instructions implementing $U^\dagger$, $V_0^\dagger \oplus V_1^\dagger$.

\subsubsection{Example of usage}
In our example, we will use $U =  H$ (the Hadamard gate).
% To keep us focused on the implementation in PyQBench and not delve too deep into mathematical explanation, we simply provide explicit formulas for discriminator $\ket{\Psi_0} $, $V_0$ and $V_1$, leaving the calculations to the interested reader.
The explicit formula for discriminator in this model reads
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with corresponding parts of optimal measurement being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} \sin\left( \frac{3}{8} \pi \right).
\end{equation}
As a next step, we need decompose our matrices into actual sequences of gates.
The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$ is well known, and comprises Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the discriminator $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = RY\left(\frac{3}{4} \pi \right) $, where $RY$ is rotation gate around the $Y$ axis defined by
\begin{equation}
RY(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, which is equivalent to following $V_0$ by $X$ matrix, more precisely
\begin{equation}
V_1 =  RY\left(\frac{3}{4} \pi \right) \cdot X.
\end{equation}
Finally, the optimal probability of correct discrimination between $\PP_{U}$, where $U=H$ and $\PP_{\Id}$ is given by 
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}. 
\end{equation}
The detailed calculation we can see in \ref{app:hadamard}.

Now we simulate  theoretical framework using PyQBench. We will start with simulation in Qiskit SDK with Aer simulator. First, we import necessary modules. \todo[inline]{czy to zdanie jest poprawne?}

\begin{lstlisting}[language=Python]
from qiskit import QuantumCircuit, Aer
import numpy as np
\end{lstlisting}
Next we import needed functions from PyQBench. For our example we will need two functions. 
\begin{lstlisting}[language=Python]
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To run any of these functions, we need to define components that we discussed below. Most importantly, we need to create Qiskit instructions, not circuits. We can conveniently do so by constructing circuit acting on qubits 0 and 1 and then converting them using \textit{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

@@

\end{lstlisting}
For the purpose of this example,  we define Qiskit Aer simulator.
\begin{lstlisting}[language=Python, caption= Defining a backend]
simulator = Aer.get_backend("aer_simulator")

@@

\subsubsection{Gaining more control over the benchmarking process}
In the example presented above we used functions that automate the whole process - from the circuit assembly, through running the simulations to interpreting the results. But what if we want more control over some parts of this process? 
One possibility would be to add more and more parameters to \textit{benchmark\_using\_xyz} functions, but this approach clearly is not very scalable. Also, there is no way we can anticipate all the possible use cases.
 We decided on another approach.  PyQBench provides functions performing:
\begin{enumerate}
\item assembly of circuits needed for experiment, provided the components discussed above
\item interpretation of the obtained measurements
\end{enumerate}
 The difference between the two approaches is illustrated on the diagrams below.

@@

Consider the particular experiment data

>>>>

\author{Paulina Lewandowska}
\author{\L ukasz Pawela}

@@

\todo[inline]{lewandowska add cite !!!}

Noisy Intermediate-Scale Quantum (NISQ)~\cite{preskill} devices are storming the market,
with a wide selection of devices based on different architectures and accompanying software
solutions. Among hardware providers offering public access to their gate--based devices, one could
mention Rigetti, IBM, Oxford Quantum Group, IonQ or Xanadu. Other vendors offer devices operating in
different paradigms. Notably, one could mention D-Wave and their quantum
annealers, or QuEra devices based on neural atoms.  Most vendors provide their own software stack and
application programming interface for accessing their devices. To name a few, Rigetti's computers
are available through their Forest SDK and PyQuil library and IBM Q computers can be accessed
through Qiskit or IBM Quantum Experience web interface. Some cloud services, like Amazon Braket, offer access to several quantum devices under a unified API. On top of that, several
libraries and frameworks can integrate with different hardware vendors. Examples of such frameworks
include IBM Q's Qiskit or Zapata Computing's Orquestra.

It is well known that NISQ devices have their limitations \cite{}. The question is to what extent those
devices can perform meaningful computations? To answer this question, one has to devise a
methodology for benchmarking them. For gate--based computers, on which this paper focuses, there
already exist several approaches. 
One could mention randomized benchmarking  \cite{knill2008randomized} , benchmarks based on the quantum volume\cite{cross2019validating, moll2018quantum} or or cross-entropy benchmarking ~\cite{arute2019quantum, neill2018blueprint}.

In this paper, we introduce yet another method for benchmarking gate--based devices with a simple
operational interpretation. In our method, we test how well the given device can distinguish between
two von Neumann measurements performed in different bases. We implemented our method in an
open-source Python library called PyQBench. The library supports any device available through Qiskit
library, and thus can be used with providers such as IBM Q or Amazon Braket. Along with the library,
the PyQBench package contains a command line tool for running most common benchmarking scenarios.

\section{Existing benchmarking methodologies and software}
Unsurprisingly, PyQBench is not the only software package for benchmarking gate--based devices.
While we believe that our approach has significant benefits over other benchmarking techniques, for
completeness, in this section we discuss some of the currently available similar software.

Probably the simplest benchmarking method one could devise is simply running known algorithms and
comparing outputs with the expected ones. Analyzing the frequency of the correct outputs, or the
deviation between actual and expected outputs distribution provides then a metric of the performance
of a given device. Libraries such as Munich Quantum Toolkit (MQT) Benchmark library contain
benchmarks leveraging multiple algorithms, such as Shor's algorithm or Grover's algorithm, 
Despite being intuitive and easily interpretable, such benchmarks may have some problems. Most
importantly, they assess the usefulness of a quantum device only for a very particular algorithm,
and it might be hard to extrapolate their results to other algorithms and applications. For
instance, the inability of a device to consistently find factorizations using Shor's algirithms
does not tell anything about its usefulness in Variational Quantum Algorithm's.

Another possible approach to benchmarking quantum computers is randomized benchmarking
\cite{liu2022sampling, knill2007randomized, wallman2014randomized, helsen2022general,
	cornelissen2021scalable}. In this approach, one samples circuits to be run from some predefined set
of gates (e.g. from the Clifford group) and tests how much the output distribution obtained from the
device running these circuits differs from the ideal one. It is also common to concatenate randomly
chosen circuits with their inverses (which should yield the identity circuit) and run those
concatenated circuits on the device. Libraries implementing this approach include Qiskit \cite{},
\todo[inline]{enumerate other libraries and summary}

@@

In general, the discrimination scheme, presented in Fig.~\ref{fig:theoretical_scheme}, requires an
auxiliary system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, the final measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise we guess that it was $\mathcal{P}_\Id$.

@@

Current NISQ devices are unable to perform conditional measurements, which is the biggest
obstacle to implementing our scheme on real hardware. However, we can adjusting our scheme so that it only uses components available on current devices.
For this purpose, we use two possible options: using a postselection or a direct sum
$V_0^\dagger\oplus V_1^\dagger$.

@@

	The first idea uses a postselection scheme. Instead of performing a conditional measurement $\PP_{V_k}$, for each
	measurement $\PP_U, \PP_\Id$ to be discriminated and each choice of $k \in \{0, 1\}$ we run circuits presented in Fig. \ref{fig:postselection}.

@@

		\caption{
			A schematic representation of the setup for distinguishing
			measurements $\PP_{U}$ and $\PP_{\Id}$ using postselection approach.
		}\label{fig:postselection}
	\end{figure}

The experiments can be grouped into classes identified by tuples of the form
	$(\mathcal{Q}, k, i, j)$, where $\mathcal{Q} \in \{\PP_U, \PP_\Id\}$ denotes the chosen measurement, $k \in \{0,1\}$, $i \in \{0,1\}$ and $j \in \{0,1\}$.
	In postselection approach, we discard all the experiments for which $k \ne i$. Hence, the total number of valid experiments is

@@

	Finally, we count the valid experiments resulting in successful discrimination.
	If we have chosen $\PP_U$, then we guess correctly iff $j=0$. Similarly, for
	$P_\Id$, we guess correctly iff $j=1$. If we define
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_U, k = i, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_\Id, k = i, j = 1\},
	\end{eqnarray}
	then the empirical success probability can be computed as

@@

The second idea uses the direct sum $V_0^\dagger \oplus V_1^\dagger$ implementation.     Here, instead of performing a conditional
measurement $\PP_{V_k}$, where $k\in \{0,1\}$,  we run circuits presented in	Fig.~\ref{fig:controlled}.

	\begin{figure}[h!]
		\centering
		\includegraphics[scale=1.5]{pics/controlled_unitary}

@@

So far, we only discussed how the discrimination is performed using two different
schemes, assuming that all needed components ($\ket{\psi_0}$, $V_0$, $V_1$) are known. It turned out that the determination of $V_0$ and $V_1$ is a simple task while the discriminator $\ket{\psi_{0}}$ is known. However,
determining the analytical form of the discriminator $\ket{\psi_{0}}$
is, in general, NP-hard problem.
We have found the analytical solution of the discrimination scheme for a particular class of von Neumann measurements.

@@

\begin{equation}
U = H
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this case, the optimal input state is the Bell state of the form
\begin{equation}
\ket{\psi_{0}} = \frac{1}{\sqrt{2}} \ket{00} + \ket{11}.
\end{equation}
Observe that the discriminator does not depend on the angle $\phi$. 
Nevertheless, for a given angle  $\phi \in  [0, 2 \pi)$,   the unitaries $V_0$,  $V_1$
have the following form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi -
	\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right)
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
Finally, we have also calculated the theoretical probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ is given by
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} .
\end{equation}
More details and proofs of the construction's validity   we can find in the \ref{app:optimal-probability} and \ref{app:optimal-strategy}.

@@

\section{PyQBench as a library}\label{sec:pyqbench-as-library}
This guide introduces basic concepts needed for using PyQBench as a library from a Python script. In particular, we will consider the following topics:
\begin{enumerate}
\item Defining experiment using circuit components.
\item Assembling circuits needed for given benchmarking scheme.
\item Running experiment on a simulator or a hardware.
\item Obtaining empirical probability of successful discrimination between measurements.
\end{enumerate}

On the conceptual level, every discrimination experiment performed in PyQBench needs the following:
\begin{enumerate}
\item Discriminator, i.e. optimal initial state for the circuit.
\item Unitary $U^\dagger$ defining alternative measurement to be discriminated from the measurement in Z-basis.
\item $V_0^\dagger$ and $V_1^\dagger$, positive and negative parts of optimal measurement from Holevo-Helstrom theorem.
\end{enumerate}

\subsubsection{Example of usage}
In  this example, we will use $U =  H$ (the Hadamard gate).
The explicit formula for discriminator in this model reads
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with corresponding parts of optimal measurement being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} \sin\left( \frac{3}{8} \pi \right).
\end{equation}
As a next step, we need decompose our matrices into actual sequences of gates.
The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$  comprises the Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the discriminator $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = RY\left(\frac{3}{4} \pi \right) $, where $RY$ is rotation gate around the $Y$ axis defined by
\begin{equation}
RY(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, which is equivalent to 
\begin{equation}
V_1 =  RY\left(\frac{3}{4} \pi \right) \cdot X.
\end{equation}
Finally, calculate the optimal probability of correct discrimination which is equal to
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}. 
\end{equation}
The detailed calculation we can see in \ref{app:hadamard}. 
Now we simulate the  theoretical framework using PyQBench. We will start with simulation in Qiskit SDK with Aer simulator. First, we import necessary modules and needed functions from PyQBench.

\begin{lstlisting}[language=Python]
import numpy as np
from qiskit import QuantumCircuit, Aer
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To run a function from PyQBench, we need to define Qiskit instructions.  We can do so by constructing circuit acting on qubits 0 and 1 and then converting them using \textit{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

@@

\end{lstlisting}
For the purpose of this example,  we run experiments on Qiskit Aer simulator.
\begin{lstlisting}[language=Python, caption= Defining a backend]
simulator = Aer.get_backend("aer_simulator")

@@

\subsubsection{Gaining more control over the benchmarking process}
In the example presented above we used functions that automate the whole process - from the circuit assembly, through running the simulations to interpreting the results. But what if we want more control over some parts of this process? 
One possibility would be to add some additional parameters to \textit{benchmark\_using\_xyz} functions, but this approach is not scalable. Moreover, there is no way we can anticipate all the possible use cases.
 We decided on another approach.  PyQBench provides functions performing:
\begin{enumerate}
\item assembly of circuits needed for experiment, provided the components discussed above
\item interpretation of the obtained measurements
\end{enumerate}
 The difference between the two approaches is illustrated on the diagrams below.

@@
